---
title: "PS270 Final Project"
author: "Laney Cardona"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Project Ideas 

I’m interested in looking at the poll data based on factors such as states and voter demographics. I think it would be interesting to see if there are any correlations or relationships between voters and their attitudes regarding certain issues, such as immigration. For example, looking to see if people who have varying political ideologies had different levels of attitude change on immigration policies compared those with a different ideology. Additionally, looking to see how their political ideologies correlate between voting patterns and their overall attitudes regarding immigration policies could also be interesting. 


## My Research Question 

How are political ideologies associated with attitudes toward immigration policies among participants, and are there differences in the level of attitude shifts regarding these policies depending on the demographic of the participant?

My hypothesis is that participants with more conservative ideologies will be less likely to support increasing the number of immigrants allowed in the country. Additionally, I think there will be a greater positive shift in attitudes towards pro-immigration policies within this group. People with more conservative beliefs and values tend to oppose increasing the amount of immigrants in the country, as they often perceive immigrants as a competition for jobs and resources. However, I think that after having an interaction with an immigrant, those who tend to favor anti-immigration policies may have a greater attitude shift towards pro-immigration policies compared to those who have more liberal ideologies. I expect that people who are more liberal will have a smaller shift after this interaction, as they already tend to favor pro-immigration policies. The explanatory variable is the political ideologies of the participants. This is measured through the survey data that represents each participant from the experiment. The outcome variable is the attitude shift regarding immigration policies between participants who have different political ideologies. This is measured by observing the attitudes of the participants before and after interacting with the confederate immigrant, to determine which group has a greater attitude shift regarding immigration policies. Patterns that show participants who were conservative were more likely to start out favoring a decrease in immigrants in the country would support my hypothesis. Additionally, my hypothesis would also be supported if after the encounter with the immigrant, conservative participants had a greater shift in positive attitudes towards immigration policies compared to those with liberal ideologies. However, if patterns show a negative correlation, in which the interaction with an immigrant resulted in either a smaller shift in attitudes for conservatives compared to liberals, or if there was a negative shift in attitudes, in which conservatives were even less likely to support immigrants in the country after the interaction, then my hypothesis would be disproven. 


```{r}
library(ps270data)
trains
```


```{r}
library(tidyverse)
trains <- trains |>
  mutate(ideology_group = case_when(
    ideology == 1 ~ "1. Most Liberal",
    ideology == 2 ~ "2. Liberal", 
    ideology == 3 ~ "3. Moderate",
    ideology == 4 ~ "4. Conservative",
    ideology == 5 ~ "5. Most Conservative"
  ))
trains
```



```{r}
average_change <- trains |>
  mutate(treatment = if_else(treatment == 1, "Treated", "Control")) |>
  filter(treatment == "Treated") |>
  group_by(ideology_group) |>
  summarize(numberim.post = mean(numberim.post, na.rm = TRUE),
            numberim.pre = mean(numberim.pre, na.rm = TRUE)) |>
  mutate(average_change = numberim.post - numberim.pre)
average_change
```
## Bar graph 
```{r}
ideology_plot <- ggplot(average_change, aes(x = ideology_group, 
                                            y = average_change)) +
  geom_bar(stat = "identity") +
   labs(title = "Average Change in Attitudes by Ideology",
       x = "Ideology Group",
       y = "Average Change in Immigration Policy Attitudes")
ideology_plot
```
## Plotted points - another graph option 
```{r}
ideology_plot <- ggplot(average_change, aes(x = ideology_group,
                                            y = average_change)) +
  geom_point(color = "purple", size = 3.5) +
  labs(title = "Average Change in Attitudes by Ideology",
       x = "Ideology Group",
       y = "Average Change in Immigration Policy Attitudes") +
  geom_smooth(method = "lm", se = FALSE, color = "darkgreen", size = 1.5) 
ideology_plot
```







































