---
title: "PS270 Final Project"
author: "Laney Cardona"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Project Ideas 

I’m interested in looking at the poll data based on factors such as states and voter demographics. I think it would be interesting to see if there are any correlations or relationships between voters and their attitudes regarding certain issues, such as immigration. For example, looking to see if people who have varying political ideologies had different levels of attitude change on immigration policies compared those with a different ideology. Additionally, looking to see how their political ideologies correlate between voting patterns and their overall attitudes regarding immigration policies could also be interesting. 


## My Research Question 

How are political ideologies associated with attitudes toward immigration policies among participants, and are there differences in the level of attitude shifts regarding these policies depending on the demographic of the participant?

My hypothesis is that participants with more conservative ideologies will be less likely to support increasing the number of immigrants allowed in the country. Additionally, I think there will be a greater positive shift in attitudes towards pro-immigration policies within this group. People with more conservative beliefs and values tend to oppose increasing the amount of immigrants in the country, as they often perceive immigrants as a competition for jobs and resources. However, I think that after having an interaction with an immigrant, those who tend to favor anti-immigration policies may have a greater attitude shift towards pro-immigration policies compared to those who have more liberal ideologies. I expect that people who are more liberal will have a smaller shift after this interaction, as they already tend to favor pro-immigration policies. The explanatory variable is the political ideologies of the participants. This is measured through the survey data that represents each participant from the experiment. The outcome variable is the attitude shift regarding immigration policies between participants who have different political ideologies. This is measured by observing the attitudes of the participants before and after interacting with the confederate immigrant, to determine which group has a greater attitude shift regarding immigration policies. Patterns that show participants who were conservative were more likely to start out favoring a decrease in immigrants in the country would support my hypothesis. Additionally, my hypothesis would also be supported if after the encounter with the immigrant, conservative participants had a greater shift in positive attitudes towards immigration policies compared to those with liberal ideologies. However, if patterns show a negative correlation, in which the interaction with an immigrant resulted in either a smaller shift in attitudes for conservatives compared to liberals, or if there was a negative shift in attitudes, in which conservatives were even less likely to support immigrants in the country after the interaction, then my hypothesis would be disproved. 


```{r}
library(ps270data)
trains
```

```{r}
library(tidyverse)
trains <- trains |>
  mutate(ideology_group = case_when(
    ideology == 1 ~ "1. Most Liberal",
    ideology == 2 ~ "2. Liberal", 
    ideology == 3 ~ "3. Moderate",
    ideology == 4 ~ "4. Conservative",
    ideology == 5 ~ "5. Most Conservative"
  ))
trains
```

```{r}
average_change <- trains |>
  mutate(treatment = if_else(treatment == 1, "Treated", "Control")) |>
  filter(treatment == "Treated") |>
  group_by(ideology_group) |>
  summarize(numberim.post = mean(numberim.post, na.rm = TRUE),
            numberim.pre = mean(numberim.pre, na.rm = TRUE)) |>
  mutate(average_change = numberim.post - numberim.pre)
average_change

trains <- trains |>
   mutate(average_change = numberim.post - numberim.pre)

```

```{r}
pre_treatment <- trains |>
  mutate(treatment = if_else(treatment == 1, "Treated", "Control")) |>
  filter(treatment == "Treated") |>
  group_by(ideology_group) |>
  summarize(numberim.pre = mean(numberim.pre, na.rm = TRUE)) |>
  mutate(pre_treatment = numberim.pre)
pre_treatment
```
```{r}
post_treatment <- trains |>
  mutate(treatment = if_else(treatment == 1, "Treated", "Control")) |>
  filter(treatment == "Treated") |>
  group_by(ideology_group) |>
  summarize(numberim.post = mean(numberim.post, na.rm = TRUE)) |>
  mutate(post_treatment = numberim.post)
post_treatment
```

## Bar Graph Showing Participant Attitudes Toward Immigration Policies Before Treatment
```{r}
pre_treatment_plot <- ggplot(pre_treatment, aes(x = ideology_group,
                                               y = pre_treatment)) +
  geom_bar(stat = "identity") +
  labs(title = "Attitudes Toward Immigration Policies by Ideology Before Treatment",
       x = "Ideology Group",
       y = "Attitudes Towards Immigration in the Country
       (1 = Increase, 5 = Decrease)",
       subtitle = "Participants Were Asked if the Number of Immigrants in the Country Should Increase or Decrease") +
  theme_minimal()
pre_treatment_plot
```

## Bar Graph Showing the Participant's Shift in Attitudes Toward Immigration Policies After Treatment 
```{r}
ideology_shift_plot <- ggplot(average_change, aes(x = ideology_group, 
                                            y = average_change)) +
  geom_bar(stat = "identity") +
   labs(title = "Shift in Attitudes Toward Immigration Policies by Ideology After Treatment",
       x = "Ideology Group",
       y = "Average Change in Immigration Policy Attitudes",
       subtitle = "The Participants' Change in Attitude After Interacting with an Immigrant") +
  theme_minimal()
ideology_shift_plot
```

## Bar Graph Showing Participant Attitudes Towards Immigration Policies After Treatment 
```{r}
post_treatment_plot <- ggplot(post_treatment, aes(x = ideology_group, 
                                                 y = post_treatment)) +
  geom_bar(stat = "identity") +
  labs(title = "Attitudes Toward Immigration Policies by Ideology After Treatment",
       x = "Ideology Group",
       y = "Attitudes Towards Immigration in the Country
       (1 = Increase, 5 = Decrease)",
       subtitle = "Participants Were Asked the Same Question After Interacting With the Immigrant") +
  theme_minimal()
post_treatment_plot
```




## Results


The table shows the average attitudes regarding immigration in the country before and after interacting with an immigrant from a scale of 1-5, where 1 represents the most pro-immigration attitudes and 5 represents the least pro-immigration attitudes. Additionally, it highlights the average change in attitudes for each ideology group due to this interaction. The table indicates that the most conservative participants were least supportive of increasing the number of immigrants in the country, while the most liberal participants were most supportive. This is to be expected, as the conservative ideology tends to be more supportive of stronger border control to limit the accessibility for immigrants to come into the country. The graph that represents the shift in attitudes regarding immigration policies after the participants interacted with the immigrant indicates that the participants in the "Most Conservative" ideology group had the greatest shift in attitudes compared to the other ideology groups. Additionally, the graph shows this ideology group had a significant negative shift, meaning that after they interacted with the immigrant, their attitudes shifted towards more pro-immigration policies, supporting my hypothesis. Finally, the last column in the table represents the attitudes of the participants after interacting with the immigrant. As I hypothesized, the attitudes regarding immigration policies after the treatment shows that the more liberal ideological groups tended to have a smaller shift in attitude. This is also consistent within liberal ideologies, as these political groups already tend to favor pro-immigration policies. Therefore, according to this data, political ideologies are associated with attitudes toward immigration policies. Furthermore, the "Most Conservative" ideology group had a much more significant shift in attitudes regarding these policies compared to the "Most Liberal" ideology group.


## Introduction

My research observed how political ideologies are associated with attitudes toward immigration policies, specifically looking how the level of attitude shifts regarding these policies differed depending on the political demographic of the participant. My hypothesis is that the participants with more conservative ideologies will be initially less likely to support increasing the number of immigrants allowed in the country, but after interacting with foreigners they will have a greater positive shift in attitudes towards pro-immigration policies compared to those with democratic views. I expect that people who are more liberal will have a smaller shift after this interaction, as they already tend to favor pro-immigration policies. 

This is significant, particularly in our country where immigration is a very politically charged topic, as it may seek to explain how individuals' beliefs and values evolve, while also bridging the current nation-wide political divide. By exploring how there may be some common ground between Conservatives and Liberals, this research may be able to help dismantle political polarization and enable greater opportunities for healthy discourse that can offer potential solutions to current policy debates, in this case regarding immigration. 


## Data

The data for this research comes from a randomized field experiment conducted by Harvard Professor, Ryan Enos. He focused on individuals from overwhelmingly white communities who rode the communal train and explored how exposure to demographic change affected their political views. His study was published in the article "Causal Effect of  Intergroup Contact on Exclusionary Attitudes." In the experiment, the trains were randomly assigned to either the "Treated group" or the "Control group," where the treatment involved the presence of two native Spanish-speaking confederates meant to simulate the predicted demographic change in the United States. Those who were in the control group never encountered the confederates.

The independent variable is the political ideologies of the participants, measured by the survey data for each participant from the experiment. Political ideologies are likely to impact individuals' pre-existing beliefs, which would set a baseline for their attitudes towards immigration before treatment, and may impact their responses to treatment. Through this method, we can determine how the treatment effects differ depending on the participants' political ideology. 

The dependent variable, the participants' shift in attitudes regarding immigration policies, is measured using a survey question, on a 5 point scale, before and after the treatment. Participants were asked whether or not the number of immigrants in the country should increase or decrease, where 1 indicates the participant believes the number of immigrants should increase and 5 indicates the participant believes the number of immigrants should decrease. The attitude shift of each ideology group is represented by the difference between their responses before and after treatment. 

```{r}
pre_treatment_plot <- ggplot(pre_treatment, aes(x = ideology_group,
                                               y = pre_treatment,
                                               fill = ideology_group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("1. Most Liberal" = "blue",
                               "2. Liberal" = "blue",
                               "4. Conservative" = "red",
                               "5. Most Conservative" = "red")) +
  labs(title = "Attitudes Toward Immigration Policies by Ideology Before Treatment",
       x = "Ideology Group",
       y = "Attitudes Towards Immigration in the Country
       (1 = Increase, 5 = Decrease)",
       subtitle = "Participants were asked if the number of immigrants in the country should increase or decrease") +
  theme_minimal() +
    theme(
    axis.text.x = element_text(size = 9, family = "Georgia"),
    axis.text.y = element_text(size = 10, family = "Georgia"),
    axis.title.x = element_text(size = 12, face = "bold", family = "Georgia"),
    axis.title.y = element_text(size = 11, face = "bold", family = "Georgia"),
    plot.title = element_text(size = 13, face = "bold", family = "Georgia"),
    plot.subtitle = element_text(size = 10, family = "Georgia"),
    legend.title = element_text(size = 11, face = "bold", family = "Georgia"),
    legend.text = element_text(size = 9, family = "Georgia"),
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.background = element_rect(color = "black", fill = NA)
  )
pre_treatment_plot
```

This table shows the relationship between the political ideologies of the participants and how they felt about whether immigrants in the country should increase or decrease before treatment. From this table, the data indicates that the participants who were most conservative had the most negative views towards immigration in the country, whereas the liberal participants had more positive views towards immigration. This is to be expected, as these beliefs align with the traditional values of Conservative and Liberal political ideologies. My research explores how the differing political ideologies will result in different levels of shifts in attitudes after recieving treatment. 

## Results

```{r}
ideology_shift_plot <- ggplot(average_change, aes(x = ideology_group, 
                                            y = average_change,
                                            fill = ideology_group)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("1. Most Liberal" = "blue",
                               "2. Liberal" = "blue",
                               "4. Conservative" = "red",
                               "5. Most Conservative" = "red")) +
   labs(title = "Shift in Attitudes Toward Immigration Policies by Ideology After Treatment",
       x = "Ideology Group",
       y = "Average Change in Immigration Policy Attitudes",
       subtitle = "The participants were asked the same question after interacting with the foreigners") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 9, family = "Georgia"),
    axis.text.y = element_text(size = 10, family = "Georgia"),
    axis.title.x = element_text(size = 12, face = "bold", family = "Georgia"),
    axis.title.y = element_text(size = 11, face = "bold", family = "Georgia"),
    plot.title = element_text(size = 12, face = "bold", family = "Georgia"),
    plot.subtitle = element_text(size = 10, family = "Georgia"),
    legend.title = element_text(size = 11, face = "bold", family = "Georgia"),
    legend.text = element_text(size = 9, family = "Georgia"),
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    legend.background = element_rect(color = "black", fill = NA)
  )
ideology_shift_plot
```
This bar graph shows the average change in attitudes towards immigration policies after treatment, representing the difference in the attitudes from before and after participants interacted with the foreigners. The average change in attitudes shows either a positive or negative shift from their pre-existing beliefs before treatment. The scale used in this analysis indicates that a value of 1 reflects a positive attitude towards immigration, while a value of 5 reflects a negative attitude towards immigration. Since lower values reveal a more positive attitude towards immigration, a decrease in value after treatment suggests a positive shift in attitudes towards immigration. On the other hand, an increase in value after treatment would suggest a negative shift in attitudes towards immigration. From this data, the participants in the "Most Conservative" group experienced a strong positive shift in attitudes after treatment, forming a more supportive stance to immigration than before. Interestingly, the participants in the "Liberal" group experienced a negative shift away from positive immigration attitudes.


```{r}
library(broom)
reg <- lm(average_change ~ ideology_group, data = trains)
reg_table <- tidy(reg)
reg_table

```


```{r}
trains <- trains |>
  mutate(binary_variable = case_when(
    ideology_group %in% c("1. Most Liberal", "2. Liberal") ~ 0,
    ideology_group %in% c("4. Conservative", "5. Most Conservative") ~ 1
  )) |>
  filter(!ideology_group %in% c("3. Moderate"))
trains
```

```{r}
reg1 <- lm(average_change ~ binary_variable, data = trains)
reg1_table <- tidy(reg1)
reg1_table
```
To create a clearer analysis that observed how liberals and conservatives as a whole differed in their shifts in attitude, I created a binary variable, in which the "Most Liberal" and "Liberal" groups were assigned to 0, and the "Conservative" and "Most Conservative" groups were assigned to 1. This regression output shows the relationship between these two combined ideology groups and their average change in attitudes towards immigration policies in the United States. The Intercept category represents the average change in attitudes for the reference group, here it refers to the liberal groups. The binary variable category compares the conservative groups to the liberal groups. The binary variable category compares the conservative groups to the liberal groups. The main coefficient for the intercept is 0.07407407, indicating that the average change in attitudes for liberals slightly increased. The main coefficient for the binary variable is -0.38441890, suggesting that in relation to the liberals, on average the conservatives experienced shifts that were about 0.384 less than liberals. Therefore, based on the scale of this experiment, the liberals experienced a slightly negative change in attitudes towards immigration policies, whereas the conservatives had a shift towards more positive attitudes towards immigration policies. 




```{r}

reg1_plot <- ggplot(reg1_table, aes(x = term, y = estimate)) +
  geom_point(size = 3, color = c("blue", "red")) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  labs(
    title = "Regression Results",
    x = "Ideology",
    y = "Coefficient"
  ) +
  scale_x_discrete(labels = c("Liberal", "Conservative")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, family = "Georgia"),
    axis.text.y = element_text(size = 12, family = "Georgia"),
    axis.title.x = element_text(size = 13, face = "bold", family = "Georgia"),
    axis.title.y = element_text(size = 13, face = "bold", family = "Georgia"),
    plot.title = element_text(size = 18, face = "bold", family = "Georgia"),
    panel.border = element_rect(color = "black", fill = NA, size = 1)
  )
  
reg1_plot
```
This graph reflects a visualization of the regression. 
